---
title: "Practice-3"
author: "Harsh"
date: "25/05/2020"
output:
  pdf_document: default
  html_document: default
---

1. Download the data set for the tutorial.
2. Follow this tutorial on applying kNN to prostate cancer detection and implement all of the steps in an R Notebook. Make sure to explain each step and what it does.

```{r}

library(data.table)

cancer_data <- read.csv("C:\\Users\\Madhav\\Desktop\\Harsh\\Rproj\\prostate_cancer.csv", header = TRUE, stringsAsFactors = FALSE)

head(cancer_data)

#Creating a copy of cancer_data
cancer_data_1 <- cancer_data

#Deleting the first column which is id 
cancer_data_1 <- cancer_data_1[-1]

#Checking the total number of patients
table(cancer_data$diagnosis_result)

setDT(cancer_data_1)

#Creating a Normalize function using min/max method. This helps in 
normalize <- function(x){
  return((x - min(x)) / (max(x) - min(x)))
}

#Normalizing the values by calling the function with the help of lapply
cancer_data_1 <- as.data.frame(lapply(cancer_data[,3:9],normalize))

summary(cancer_data_1)

#Dividing the data in 65:35 ratio
training_data <- cancer_data_1[1:65,]
testing_data <- cancer_data_1[66:100,]

#Selecting the labels and storing in a new variable
training_data_labels <- cancer_data[1:65,2]
testing_data_labels <- cancer_data[66:100,2]

```


```{r}

#install.packages("class")
#install.packages("gmodels")

library(class)
library(gmodels)

#Using the knn function from the class library. We select the efficient value which is k = 7. We get an accuracy of 71.4% with k = 7 and With k = 10 we get 62%
cancer_pred_1 <- knn(train = training_data, test = testing_data, cl = training_data_labels, k = 10)
cancer_pred <- knn(train = training_data, test = testing_data, cl = training_data_labels, k = 7)

CrossTable(x = testing_data_labels, y = cancer_pred_1, prop.chisq = FALSE)
CrossTable(x = testing_data_labels, y = cancer_pred, prop.chisq = FALSE)

```
We get an accuracy of around 60% for k = 10 which is quite low. We try to improve the accuracy by testing with different k values, the best one achieved is for k = 7 i.e 71%.


3. Once you've complete the tutorial, try another kNN implementation from another package, such as the caret package. Compare the accuracy of the two implementations.
4. Try the confusionMatrix function from the caret package to determine the accuracy of both algorithms.

```{r}

#install.packages("caret")
library(caret)

cancer_data$diagnosis_result <- as.factor(cancer_data$diagnosis_result)

set.seed(123)

#Creating new data variables for testing caret package
train.cancer <- training_data
train.cancer$diagnosis_result <- as.factor(training_data_labels)
test.cancer <- testing_data
test.cancer$diagnosis_result <- as.factor(testing_data_labels)

#Assigning trainControl values to a variable
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

#Using train function from the caret package to train the model
knn_fit <- train(diagnosis_result ~., data = train.cancer, method = "knn",
                 trControl=trctrl,
                 preProcess = c("center", "scale"),
                 tuneLength = 10)

#Calling the trained function to get the accuracy of different k values
knn_fit

#Plotting accuracy vs k values
plot(knn_fit)

#Testing the model with 35% test data
test_pred <- predict(knn_fit, newdata = test.cancer)
test_pred

# We get an accuracy of 68.57% for Caret Package
confusionMatrix(test_pred, test.cancer$diagnosis_result)

# We get an accuracy of 71.43% for Class Package
confusionMatrix(cancer_pred, as.factor(testing_data_labels))

```

Results of training model are very high for k = 15 which is 88% accuracy but when we test the model with testing data, observed accuracy is very low which is 68% for caret package compared to class package which is 71%.
