---
title: "Practicum-3.2"
author: "Harsh"
date: "26/07/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r}

#Importing Libraries
library(fpc)

library(arules)
#unloadNamespace("arules") 
#update.packages("arules")
#library(arules)

```

                                          ---Problem 2---

1. Download the data set Plant Disease Data Set. Note that the data file does not contain header names; you may wish to add those. The description of each column can be found in the data set explanation. This assignment must be completed within a separate R Markdown Notebook. Use read.transaction() from the arules package to read the data.
2. Explore the data set as you see fit and that allows you to get a sense of the data and get comfortable with it. Is there distributional skew in any of the features? Is there a need to apply a transform?

- I used read.transactions function to import the plants transaction data
- Since this the data is transaction i.e sparse matrix we cannot observe the skewness or distribution of the data
- Using summary and inspect I did explanatory analysis

```{r}

#Importing Dataset
plant_data <- read.transactions("C:\\Users\\harsh\\Desktop\\Introduction to Machine learning and Data Mining\\Practicum-3\\plants.data",sep = ",")

#Observing transactions
summary(plant_data)
inspect(plant_data[1:5])

#Checking top 20 items using itemfrequency plot
itemFrequency(plant_data[,1:5])
itemFrequencyPlot(plant_data,topN=20)

#Observing the image of the transaction data with and without sampling
image(plant_data[1:15000])
image(sample(plant_data[1:15000]))

```

3. Use association rules to segment the data similar to what was done in H채m채l채inen, W., & Nyk채nen, M. (2008, December). Efficient discovery of statistically significant association rules. In Data Mining, 2008. ICDM'08. Eighth IEEE International Conference on (pp. 203-212). IEEE.

- Using apriori function to build association rules for plants dataset
- Improving the model by adjusting the support and confidence 
- Using Inspect and summary, we can observe the output

```{r}

#using apriori function to build rules
default_rules <- apriori(plant_data)
summary(default_rules)
inspect(default_rules[1:5])

#Improving rules by changing default parameters
new_rules <- apriori(plant_data, parameter = list (support = 0.001, confidence = 0.75))
summary(new_rules)
inspect(new_rules[1:5])

```

4. Are there clusters in the data? Can plants be segmented into groups? Build a k-means clustering model to investigate. 

- First I converted the transactions to separate items and transaction ID by importing it again
- Later I stored this data in a dataframe in matrix format
- Using kmeans function, I built clusters of the data by using k = 10 for creating 10 clusters
- Using size and centers we can observe the center and size of the cluster
- Using plotCluster function I visualize the clusters
- For this problem, I referred a few links I have added them in this chunk

```{r}

#Importing the data as separate columns
plants <- read.transactions("C:\\Users\\harsh\\Desktop\\Introduction to Machine learning and Data Mining\\Practicum-3\\plants.data", sep = ",", cols = 1)

#Observing the new data
summary(plants)

#Converting the data to matrix format and as integer datatype
plants_matrix <- as.data.frame(as(plants,"matrix"))
plants_matrix[,1:70] <- lapply(plants_matrix[,1:70],as.integer)

#Verifying the data
str(plants_matrix)

#Creating the clusters using kmeans function
clusters <- kmeans(plants_matrix,10)
clusters$size
clusters$centers

#Visualizing the clusters
plotcluster(plants_matrix,clusters$cluster)

```
Reference: 
https://stats.stackexchange.com/questions/31083/how-to-produce-a-pretty-plot-of-the-results-of-k-means-cluster-analysis
https://stackoverflow.com/questions/41972270/how-to-convert-object-of-transaction-to-dataframe-in-r






